---
output: html_document
---
Brooklyn real estate sales analysis by Liya Naumova
========================================================

```{r echo=FALSE, message=FALSE, warning=FALSE, packages}

library(ggplot2)
library(openxlsx)
library(gdata)
library(dplyr)
library(stats)

```

```{r global_options,echo=FALSE}
knitr::opts_chunk$set(warning=FALSE, message=FALSE)
```

```{r echo=FALSE, Load_the_Data}
# load the data 
sales <- read.xlsx("rollingsales_brooklyn.xlsx", sheet = 1, startRow = 5,
                   detectDates = TRUE)

# convert all column names to lower case
colnames(sales) <- tolower(colnames(sales))

# substitute empty strings with NA
sales[sales$tax.class.at.present =='  ', ]$tax.class.at.present <- NA
sales[sales$building.class.at.present =='  ', ]$building.class.at.present <-NA 
sales[sales$apartment.number == "            ", ]$apartment.number <-NA

# substitute zero zipcodes with NA
sales[sales$zip.code == 0,11] <-NA

# convert all categorical values to factor
sales$neighborhood <- as.factor(sales$neighborhood)
sales$building.class.category <- as.factor(sales$building.class.category)
sales$tax.class.at.present <- as.factor(sales$tax.class.at.present)
sales$building.class.at.present <- as.factor(sales$building.class.at.present)
sales$tax.class.at.time.of.sale <- as.factor(sales$tax.class.at.time.of.sale)
sales$building.class.at.time.of.sale <-
  as.factor(sales$building.class.at.time.of.sale)
sales$block <- as.factor(sales$block)
sales$lot <- as.factor(sales$lot)
sales$zip.code <- as.factor(sales$zip.code)

# substitute zeros and probably erroneus years with NA
sales[!is.na(sales$year.built) & sales$year.built == 1380,17] <- NA
sales[!is.na(sales$year.built) & sales$year.built == 1050,17] <- NA
sales[!is.na(sales$year.built) & sales$year.built == 0, 17] <- NA

# trim whitespaces
sales$neighborhood <- trim(sales$neighborhood)
sales$building.class.category <- trim(sales$building.class.category)
sales$address <- trim(sales$address)
sales$apartment.number <- trim(sales$apartment.number)
```

I've chosen data about Brooklyn real estate sales in 2015. It consist of more than 23 000  observations of 21 variables.

```{r echo=FALSE}
str(sales)
head(sales)

```


# Univariate Plots Section
#### Price
At first I decided to look at price distribution

```{r echo=FALSE, price_summary}
summary(sales$sale.price)

```
It looks like we have zero values for about a quarter of rows and now I want to look at lowest values

```{r echo=FALSE, price_table}

head(table(sales$sale.price))

```
Then I exclude zeros and repeat summary
```{r echo=FALSE}
nonzero_prices <- subset(sales, sale.price > 0)
summary(nonzero_prices$sale.price)
```

Now I'm interested in those $1 properties

```{r echo=FALSE}
one_dollar_sale <- sales %>%
  filter(sale.price == 1) %>%
  select(building.class.category,address, year.built,
         gross.square.feet, sale.price)
head(one_dollar_sale,10)
```
Other values look normal so I assume it's a kind of fictitious price

Let's look at histogram with prices divided by 1000

```{r echo=FALSE}
ggplot(sales, aes(sale.price/1000)) +
  geom_histogram(binwidth = 1000)

```

So there is a distribution with very long tale.
Let's look at all prices lower than 5 mln.

```{r echo=FALSE}
ggplot(sales, aes(sale.price/1000)) +
  geom_histogram(binwidth = 50) +
  xlim(0.001,5000)
```

Most prices are distributed between 1000 and 1 000 000  with peaks around 500 000, 950 000 and 1 250 000, also we have much more prices that are slightly lower then 1 mln than that are slightly higher.
After log transformation prices look nearly normal but some exceptions that are lower than 1000.

```{r echo=FALSE, Price}

ggplot(sales, aes(x = sale.price/1000)) + 
  geom_histogram(bins = 100) +
  scale_x_log10()
```

#### Gross sqare feet

Next feature of interest for me is Gross square feet.

```{r echo=FALSE}
summary(sales$gross.square.feet)

```
Looks like here are also lots of zeros. Summary without zeros:
```{r echo=FALSE}
summary(subset(sales, gross.square.feet > 0)$gross.square.feet)
```


Let's look at distribution of values greater than 0

```{r echo=FALSE}
ggplot(sales, aes(gross.square.feet)) +
  geom_histogram(bins = 500) +
  scale_x_continuous(limits = c(1, 366000))

```

Here also we have very long tail. I focus on values between 1 and 10000

```{r echo=FALSE}

ggplot(sales, aes(gross.square.feet)) +
  geom_histogram(bins = 100) +
  scale_x_continuous(limits = c(1, 10000), breaks = seq(0,10000,1000))

```

Most values are between 1000 and 3500 square feet and peak is around 2000.
Now I want to look at it after log transformation.

```{r echo=FALSE, gsf_log}

ggplot(sales, aes(gross.square.feet)) +
  geom_histogram(bins = 100) + 
  scale_x_log10()

```

Now it looks more like normal but still with long tails.

#### Land sqare feet

Summary:
```{r echo=FALSE}
summary(sales$land.square.feet)
```

Looks like here we also have lots of zeros and outliers.

Summary without zeros:

```{r echo=FALSE}
summary(subset(sales, land.square.feet > 0)$land.square.feet)

```

Histogram:
```{r echo=FALSE}
ggplot(sales, aes(land.square.feet)) +
  geom_histogram(bins = 100) +
  scale_x_continuous(limits = c(1, 25000))

```

And a closer look to values between 1 and 10000:

```{r echo=FALSE}
ggplot(sales, aes(land.square.feet)) +
  geom_histogram(bins = 200) +
  scale_x_continuous(limits = c(1, 10000), breaks = seq(0,10000,1000))

```

Most values are distributed between 1500 and 3000 with spikes at round numbers (2000, 2500, 3000 ...). The most common value is 2000.

After log transformation:

```{r echo=FALSE}
ggplot(sales, aes(land.square.feet)) +
  geom_histogram(bins = 100) + 
  scale_x_log10()

```

I decided to make new variable "total.square.feet" as a sum of "land.square.feet" and "gross.square.feet".

```{r echo=FALSE}
sales <- mutate(sales, total.square.feet = land.square.feet + gross.square.feet)
```

#### Total square feet

summary of new variable:

```{r echo=FALSE}
summary(sales$total.square.feet)
```
I still have `r sum(sales$total.square.feet == 0)` zero values.
Summary of non-zero values:

```{r echo=FALSE}
summary(subset(sales, total.square.feet > 0)$total.square.feet)
```

```{r echo=FALSE}
ggplot(sales, aes(total.square.feet)) + 
  geom_histogram(bins = 100)
```
 
Obviously I have here similar distribution with very long tale.

```{r  echo=FALSE}
ggplot(sales, aes(total.square.feet)) + 
  geom_histogram(bins = 100) + 
  scale_x_continuous(limits = c(1,20000))
``` 
 
#### Price per square feet

Then I want to make another variable - price per square foot

```{r echo=FALSE}
sales$price.square.foot <- 
   ifelse(sales$total.square.feet > 0,
          round(sales$sale.price / sales$total.square.feet, 2), 0)
```

Summary of non-zero values:

```{r echo=FALSE}
summary(subset(sales, price.square.foot > 0)$price.square.foot)

```

distribution of :

```{r echo=FALSE}
ggplot(sales, aes(price.square.foot)) + 
  geom_histogram(binwidth = 50) +
  scale_x_continuous(limits = c(0, 12060))
```

distribution of non-zero values below 2000:

```{r echo=FALSE}
ggplot(sales, aes(price.square.foot)) + 
  geom_histogram(binwidth = 10) +
  scale_x_continuous(limits = c(0.01, 2000))
```

#### Year built

Next feature of interest is year.built

```{r echo=FALSE}
summary(sales$year.built)
```

It's distribution:

```{r echo=FALSE}
ggplot(sales, aes(year.built)) +
  geom_histogram(binwidth = 5) + 
  scale_x_continuous(limits = c(1800,2020), breaks = seq(1800,2020,20))
```

Most buildings were built between 1895 and 1935 with other peaks in 1950 - 1956 and 2005-2015

```{r echo=FALSE}
ggplot(sales, aes(year.built)) +
  geom_histogram(binwidth = 1) + 
  scale_x_continuous(limits = c(1800,2020), breaks = seq(1800,2020,20))
```

With binwidth = 1 it is possible to notice spikes on round years (1900, 1920...). I assume some of these values are approximate. 

#### Residential units

Table of values' counts:

```{r echo=FALSE}
table(sales$residential.units)
```

Histogram:

```{r echo=FALSE}
ggplot(sales, aes(residential.units)) + 
  geom_histogram(binwidth = 1)
```

```{r echo=FALSE}
ggplot(sales, aes(residential.units)) + 
  geom_histogram(binwidth = 1) + 
  scale_x_continuous(limits = c(0,50))
```

Most properties have from 0 to 5 residential units.

#### Commercial units

```{r echo=FALSE}
table(sales$commercial.units)
```

Histogram:

```{r echo=FALSE}
ggplot(sales, aes(commercial.units)) + 
  geom_histogram(binwidth = 1)
```

```{r echo=FALSE}
ggplot(sales, aes(commercial.units)) + 
  geom_histogram(binwidth = 1) + 
  scale_x_continuous(limits = c(0,20))
```

Most properties have 0 commercial units.

#### Total units

```{r echo=FALSE}
table(sales$total.units)
```

Histogram:

```{r echo=FALSE}
ggplot(sales, aes(total.units)) + 
  geom_histogram(binwidth = 1)
```

```{r echo=FALSE}
ggplot(sales, aes(total.units)) + 
  geom_histogram(binwidth = 1) + 
  scale_x_continuous(limits = c(0,20))
```

Most properties have between 0 and 5 total units.

### Categorical variables

#### Borough

```{r echo=FALSE}
table(sales$borough)
```
All observarions of this variable have the same value

#### Neighborhood

Categorical variable with 60 levels

```{r echo=FALSE,fig.height=8}
ggplot(sales,
       aes(x= reorder(neighborhood,neighborhood,function(x) length(x)))) + 
  geom_bar() + coord_flip() +
  labs(x = "Neighborhood")

```

I've noticed that neighborhoods have different representation in the dataset. Some of them have small number of observations.

#### Building class category

Categorical with 44 levels.
Ten most common values:

```{r echo=FALSE}
sort(table(sales$building.class.category), decreasing = TRUE)[1:10]

```

```{r echo=FALSE,fig.height=8}
ggplot(sales, aes(x = reorder(building.class.category,
                              building.class.category,
                              function(x) length(x)))) + 
  geom_bar() +
  coord_flip() +
  scale_y_log10(breaks = c(10,100,500,1000,5000)) +
  labs(x= "Building class category")
  
```

Most properties are one-three family dwellings or condos.

#### Tax class at present

Categorical with 10 levels

```{r echo=FALSE}
table(sales$tax.class.at.present)
```

#### Block

I transformed this to factor which has 5525 levels . Most frequent values are:

```{r echo=FALSE}
sort(table(sales$block), decreasing = TRUE) [1:10]
```

#### Lot

I transformed it to factor too and received 1083 different values. Most frequent of them are:

```{r echo=FALSE}
sort(table(sales$lot), decreasing = TRUE) [1:10]
```

#### Easement

```{r echo=FALSE}
table(sales$`ease-ment`)
```
All observations have empty values.

#### Building class at present

Factor with 128 levels. Most common values:

```{r echo=FALSE}
sort(table(sales$building.class.at.present), decreasing = TRUE)[1:10]
```

All values:

```{r echo=FALSE}
table(sales$building.class.at.present)
```

#### Address and apartment number

Character values that should represent uniqe buildings or apartments, I want to see if any of them repeat

```{r echo=FALSE}
sort(table(sales$address), decreasing = TRUE)[1:5]
```

#### ZIP code

factor with 39 levels that can represent geographical location of building

```{r echo=FALSE,fig.height=8}
ggplot(sales, aes(x= reorder(zip.code,zip.code,function(x) length(x)))) +
  geom_bar() + coord_flip() +
  labs(x = "zip code")
```

#### Tax class at time of sale

Factor variable with 4 levels

```{r echo= FALSE}
table(sales$tax.class.at.time.of.sale)
```
 Most values are of class 1
 
#### Building class at time of sale

Factor variable with 132 levels. Most frequent:
```{r echo=FALSE}
sort(table(sales$building.class.at.time.of.sale), decreasing = TRUE)
```

#### Sale date
I should have sales for the whole 2015 year. 

```{r echo=FALSE}
ggplot(sales, aes(sale.date)) + geom_histogram(binwidth = 1)

```

```{r echo=FALSE}
ggplot(sales, aes(sale.date)) +
  geom_histogram(binwidth =7) +
  scale_x_date(date_breaks = '1 month') +
  theme(axis.text.x = element_text(angle=60, hjust=1))

```

Obviously most sales are made on week days with spike on June, 30 and a decline to the end of year.

# Univariate Analysis


### What is the structure of your dataset?
There are 23223 property sales with 23 variables in this dataset. 8157 rows don't have information about price. For others mean price is $1237000 and mean is $660000. Most properties are between 1000 and 3500 square feet and have between 1500 and 3000 square feet of land. Most buildings were constructed between 1895 and 1935.


### What is/are the main feature(s) of interest in your dataset?
Main features for me are price, gross square feet, land square feet.

### What other features in the dataset do you think will help support your investigation into your feature(s) of interest?
Other features that I find interesting are year built, numbber of units, sale date, zipcode.

### Did you create any new variables from existing variables in the dataset?
I created a variable "total square feet" for sum of gross and land square feet and a variable for price per square feet.

### Of the features you investigated, were there any unusual distributions? Did you perform any operations on the data to tidy, adjust, or change the form of the data? If so, why did you do this?
I've deleted zeros and obvious mistakes from "year.built", for plots I deleted price by 1000 and made logarithmic transformations for price and area variables because of long-tailed distributions. 


# Bivariate Plots Section
#### Price and gross square feet
First I want to plot price vs gross square feet
```{r echo=FALSE,warning=FALSE, Bivariate_Plots}
ggplot(sales, aes(gross.square.feet, sale.price/1000)) +
  geom_point(alpha = 0.1) +
  scale_x_continuous(limits = c(0, 50000)) + 
  scale_y_continuous(limits = c(0, 50000)) 
```

Filter out zero values and zoom in:

```{r echo=FALSE,warning=FALSE}
ggplot(sales, aes(gross.square.feet, sale.price/1000)) +
  geom_point(alpha = 0.1) +
  scale_x_continuous(limits = c(1, 10000)) + 
  scale_y_continuous(limits = c(1, 10000)) + 
  geom_smooth()

```

Here we can see some vertical bands for round numbers and a horizontal stripe around $1000000 and also a lot of variance in price for the same square feet. Smothing layer shows that at average larger properties cost more. I want to calculate the correlation coeffitient:

```{r echo=FALSE,warning=FALSE}
cor = cor.test(sales$sale.price, sales$gross.square.feet)
cor$estimate
```

There is positive correlation, but not very strong.

#### Price vs. land square feet


```{r echo=FALSE,warning=FALSE}
ggplot(sales, aes(land.square.feet, sale.price/1000)) +
  geom_point(alpha = 0.1) 

```

I delete zero values and look closer 

```{r echo=FALSE,warning=FALSE}
ggplot(sales, aes(land.square.feet, sale.price/1000)) +
  geom_point(alpha = 0.1) +
  scale_x_continuous(limits = c(1, 10000)) + 
  scale_y_continuous(limits = c(1, 10000)) 
```

Here also we can see strong vertical bands and a lot of differences in prices for the same square feet.

```{r echo=FALSE,warning=FALSE}
cor = cor.test(sales$sale.price, sales$land.square.feet)
cor$estimate
```

There is low positive correlation.

#### Price vs. total square feet

```{r echo=FALSE,warning=FALSE}
ggplot(sales, aes(total.square.feet, sale.price/1000)) +
  geom_point(alpha = 0.1) +
  scale_x_continuous(limits = c(0, 50000)) + 
  scale_y_continuous(limits = c(0, 50000)) 
```

Now I filter out zero values and add linear regression line:

```{r echo=FALSE,warning=FALSE}
ggplot(sales, aes(total.square.feet, sale.price/1000)) +
  geom_point(alpha = 0.1) +
  scale_x_continuous(limits = c(1, 10000)) + 
  scale_y_continuous(limits = c(1, 10000)) + 
  geom_smooth(method = "lm")
```

As expected larger properties at average cost more, but there are also a lot of variance due to other variables.

Corellation coefficient for total square feet and price:

```{r echo=FALSE,warning=FALSE}
cor = cor.test(sales$sale.price, sales$total.square.feet)
cor$estimate
```

Correlation is lower than 0.5  suggesting that variables are correlated but not very strong.

#### Price per square foot vs. total square feet

Now I'm interested in comparing price per square foot vs. total sqare feet, may be there is a difference in price for small or large properties.

```{r echo=FALSE,warning=FALSE}
ggplot(sales, aes(total.square.feet, price.square.foot)) + 
  geom_point(alpha = 1/20) + 
  scale_x_continuous(limits = c(1,10000)) +
  scale_y_continuous(limits = c(1,1000)) + 
  geom_smooth()
```

Plot shows a lot of variance in price but no obvious increase or decrease.

#### Price per foot vs  year.built

```{r echo=FALSE,warning=FALSE}
ggplot(sales, aes(year.built, price.square.foot)) + 
  geom_point(alpha = 1/20)  
```

```{r echo=FALSE,warning=FALSE}
ggplot(sales, aes(year.built, price.square.foot)) + 
  geom_jitter(alpha = 1/30, width = 1) + 
  scale_x_continuous(limits = c(1895,2016), breaks = seq(1890,2020,20)) + 
  scale_y_continuous(limits = c(1,2000))
```

I can see more variance in prices for housed built in years 1899-1931 and 2000-2015 y but I'm not sure about changes in mean prices so I want to cut years by decades and make a boxplot.

```{r echo=FALSE}
sales$year.built.bucket <- cut(sales$year.built,
                               breaks = seq(1890,2020,10),
                               dig.lab = 4)
```

```{r echo=FALSE,warning=FALSE}
ggplot(subset(sales,price.square.foot > 0),
       aes( year.built.bucket,price.square.foot)) + 
  geom_boxplot() + 
  scale_y_continuous(limits = c(0,2000)) +
  theme(axis.text.x = element_text(angle=60, hjust=1))
```

There are small differences in median price for different decades. On average properties that were built in the beginnig of this and last century cost more than those built in the middle of last century.

#### Price per square foot vs. neighborhood

Obviously, a lot of difference in price per square foot is explaned by location.
I want to make a boxplot after filtering out zero prices and zoom on prices lower than $2500:

```{r echo=FALSE,fig.height=8}
ggplot(filter(sales, price.square.foot > 0),
       aes(neighborhood, price.square.foot)) + 
  geom_boxplot() + 
  coord_flip(ylim = c(1,2500)) 
```

#### Sale price per unit vs. neighborhood

Also we can see a lot of difference in sale prices  per unit

```{r echo=FALSE,fig.height=8,warning=FALSE}
ggplot(filter(sales, sale.price > 0),
       aes(neighborhood, sale.price/(total.units * 1000))) + 
  geom_boxplot() + 
  coord_flip(ylim = c(0,5000))
```

#### Gross square feet vs. neighborhood

Next I want to look at distribution of floor area across neighborhoods:

```{r echo=FALSE,fig.height=8, warning=FALSE}
ggplot(filter(sales,gross.square.feet >0),
       aes(neighborhood, gross.square.feet)) + 
  geom_boxplot() + 
  coord_flip(ylim = c(0,30000))
```

For example, properties that are sold  in Downtown - Fulton mall area are on average larger than in Windsor Terrace. But if I look at area per unit,  differences are not so large: 

```{r echo=FALSE,fig.height=8, warning=FALSE}
ggplot(filter(sales,gross.square.feet >0),
       aes(neighborhood, gross.square.feet/total.units)) + 
  geom_boxplot() + 
  coord_flip(ylim = c(0,30000))
```

I want to look closer to see difference:

```{r echo=FALSE,fig.height=8, warning=FALSE}
ggplot(filter(sales,gross.square.feet >0),
       aes(neighborhood, gross.square.feet/total.units)) + 
  geom_boxplot() + 
  coord_flip(ylim = c(0,10000))
```

#### Sale price vs. tax class

Now I want to look at for distribution of prices for different tax classes:

```{r echo=FALSE,warning=FALSE}
ggplot(filter(sales,sale.price > 0),
       aes(tax.class.at.present, sale.price/1000)) + 
  geom_boxplot() +
  coord_cartesian(ylim = c(0,10000))
```

#### Price per square foot vs. tax class

Next - prices per square foot:

```{r echo=FALSE, warning=FALSE}
ggplot(filter(sales,price.square.foot >0), aes(price.square.foot)) + 
  geom_density(aes(color = tax.class.at.present)) +
  coord_cartesian(xlim = c(0,2000)) +
  scale_color_brewer(palette = "Dark2")
```

Distribution for class 2C looks rather different from others

```{r echo=FALSE}
ggplot(filter(sales,price.square.foot > 0),
       aes(tax.class.at.present, price.square.foot)) + 
  geom_boxplot() +
  coord_cartesian(ylim = c(0,3000))
```

Median price for class 2C is conciderably higher then others.

####Tax class vs. Neighborhood

```{r echo=FALSE, fig.height=8}
ggplot(sales, aes(neighborhood) ) +
  geom_bar(aes(fill= tax.class.at.time.of.sale), position = "dodge") + 
  coord_flip() + scale_fill_brewer(palette = "Set1")
```

Most small residential properties (class 1) were sold in Bedford-Stuyvesant, condos (class 2) - in Park Slope, commercial (class 4) - in Bedford-Stuyvesant and Williamsburg-North. Next I want to look at proportions of different tax classes.


```{r echo=FALSE, fig.height=8}
ggplot(sales, aes(neighborhood) ) +
  geom_bar(aes(fill= tax.class.at.time.of.sale), position = "fill") + 
  coord_flip() +
  scale_fill_brewer(palette = 'Set1')
```

####Building category vs. price per foot

```{r echo=FALSE, fig.height=8,warning=FALSE}
ggplot(sales, aes(building.class.category, price.square.foot)) +
  geom_boxplot() + 
  coord_flip() + 
  scale_y_continuous(limits = c(1,1000))
```

There is some difference in median prices for diffferent building classes. For example indoor public and cultural facilities cost per foot more then educational facilities

####Sale price vs. sale date

```{r echo=FALSE,warning=FALSE}
ggplot(sales, aes(sale.date, sale.price/1000)) +
  geom_point(alpha = 1/5) +
  scale_x_date(date_breaks = '1 month') + 
  scale_y_continuous(limits = c(0,50000)) +
  theme(axis.text.x = element_text(angle=60, hjust=1))
```

Among big sales there are stripes around specific dates like 2015.03.01 or 2015.06.30.

```{r echo=FALSE,warning=FALSE}
ggplot(sales, aes(sale.date, sale.price/1000)) + 
  geom_point(alpha = 1/20) + 
  geom_line(color = 'red',stat = "summary", fun.y = mean) + 
  scale_x_date(date_breaks = '1 month') +
  scale_y_continuous(limits = c(1,10000)) +
  theme(axis.text.x = element_text(angle=60, hjust=1))
```

Red line shows mean price for every day of year.
Now I want to look how price per foot changes with time

```{r echo=FALSE,warning=FALSE}
ggplot(sales, aes(sale.date, price.square.foot)) +
  geom_point(alpha = 1/5) +
  scale_x_date(date_breaks = '1 month') + 
  scale_y_continuous(limits = c(1,1000)) +
  geom_smooth(method = 'gam') +
  theme(axis.text.x = element_text(angle=60, hjust=1))
```

There is a slight increase to the end of year.

# Bivariate Analysis

### Talk about some of the relationships you observed in this part of the investigation. How did the feature(s) of interest vary with other features in the dataset?

There is positive correlation between price and gross and land square feet, obviously larger properties cost more, but it is not so stront to explaine all the variance in price. 
Houses built in the beginning and in the end of last century at average have higher prices per square foot than those built in the middle of the century.
Price per square foot varies significantly in different neighborhoods.
For tax class 2C price distribution looks different then others, has higer mean and variance.
There is some difference in median prices for diffferent building classes. For example indoor public and cultural facilities cost per foot more then educational facilities

### Did you observe any interesting relationships between the other features (not the main feature(s) of interest)?

There is some difference in median prices for diffferent building classes. For example indoor public and cultural facilities cost per foot more then educational facilities.
Higest number of small residential properties (class 1) were sold in Bedford-Stuyvesant, condos (class 2) - in Park Slope, commercial (class 4) - in Bedford-Stuyvesant and Williamsburg-North.

### What was the strongest relationship you found?
Sale price positively corellates with floor and land area, also price is strongly related to neighborhood.

# Multivariate Plots Section

#### Price, floor area & tax class

```{r echo=FALSE,warning=FALSE}
ggplot(sales, aes(gross.square.feet, sale.price/1000,
                  color = tax.class.at.time.of.sale)) +
  geom_point(alpha = 0.3) +
  scale_x_continuous(limits = c(1, 10000)) + 
  scale_y_continuous(limits = c(1, 10000)) +
  scale_color_brewer(palette = 'Set2')
```

Tax class 1 points are mostly situated in lower left corner, class 2 in lower middle and class 3 are disperced around.
Now I,m interested in floor area distribution of different tax classes: 

```{r echo=FALSE,warning=FALSE}
ggplot(filter(sales,gross.square.feet > 0), aes(gross.square.feet)) + 
  geom_density(aes(color = tax.class.at.time.of.sale)) + 
  coord_cartesian(xlim = c(1,20000))
```

Properties larger than 4000 feet are mostly commercial or condo, and most small residentials have area less than 5000 feet.

Next I divided price and square footage by number of units and used a log scale to see relation beetween price and size of one unit across tax classes.

```{r echo=FALSE,warning=FALSE}
ggplot(filter(sales,gross.square.feet > 0,sale.price > 0),
       aes(gross.square.feet/total.units,
           sale.price/total.units)) +
  geom_point(aes(color = tax.class.at.time.of.sale),alpha = 0.3) +
  scale_y_log10() +
  scale_x_log10() +
  scale_color_brewer(palette = 'Set2') +
  geom_smooth()
```

First I noticed a group of points with prices lower than $1000, this looks strange to me, maybe these are mistakes. 
Looks like units of class 2 at average are cheaper and smaller than class 1 and units of class 4 are larger and more expensive.
I wonder does price rise at the same rate as floor area or maybe in large properties one square foot cost less?
I plotted price for square foot against floor area for unit in log scale:

```{r echo=FALSE,warning=FALSE}
ggplot(filter(sales,gross.square.feet > 0,price.square.foot > 10),
       aes(gross.square.feet/total.units,
           price.square.foot)) +
  geom_point(aes(color = tax.class.at.time.of.sale),alpha = 0.3) +
  scale_y_log10() +
  scale_x_log10() +
  scale_color_brewer(palette = 'Set2') +
  geom_smooth(aes(color = tax.class.at.time.of.sale), se = FALSE)
```

There is some evidence of downward trend for tax class 2 and bigger commercial properties.

#### Price, floor area and year built

Now I'm interested if price and square footage depends on the year.

```{r echo=FALSE,warning=FALSE}
ggplot(filter(sales,gross.square.feet > 0,sale.price > 1000, year.built >0),
       aes(gross.square.feet/total.units,
           sale.price/total.units)) +
  geom_point(aes(color = year.built),
             alpha = 0.3) +
  scale_y_log10() +
  scale_colour_gradientn(colours = heat.colors(14)) +
  coord_cartesian(xlim = c(10,3000), ylim = c(1000,10000000))
```

Looks like newer buildigs are slightly cheaper. 

#### Price per unit, tax class , year built

```{r echo=FALSE,warning=FALSE, Multivariate_Plots}
ggplot(filter(sales, sale.price >1000),
       aes(year.built.bucket,sale.price/total.units)) +
  geom_jitter(aes(color = tax.class.at.time.of.sale),
              width = 0.9, alpha = 0.2) +
  scale_y_log10() +
  theme(axis.text.x = element_text(angle=60, hjust=1))
```

Here I can notice that among the properties built in the beginning of the last century commercial (class 4) properties have higher prices than small residential (class 1) and condos (class 2). On the other hand, commercial properties built in 21st century mostly cost less than residential and condos. 

#### Price per square foot, tax class , year built

```{r echo=FALSE,warning=FALSE}
ggplot(filter(sales, price.square.foot >0),
       aes(year.built.bucket,price.square.foot)) +
  geom_jitter(aes(color = tax.class.at.time.of.sale),
              width = 0.9, alpha = 0.2) +
  coord_cartesian(ylim = c(0,2000)) +
  theme(axis.text.x = element_text(angle=60, hjust=1))
```

If I look at price per square foot I see no obvious pattern.

#### Price, tax class and neighborhood

Now I' m interested in distribution of prices for unit across neighborhoods clored by tax class.

```{r echo=FALSE,warning=FALSE,fig.height=8}
ggplot(filter(sales, sale.price > 1000),
       aes(sale.price/(total.units*1000),neighborhood)) +
  geom_jitter(aes(color = tax.class.at.time.of.sale),
              height = 0.9, alpha = 1, shape = ".") + 
  coord_cartesian(xlim = c(0,5000))
```

Here we can see that distribution varies significantly in different neighborhoods, for example, I notice clusters of lower price class 4 points in Bedford Stuyvesant and Williamsbourg and more expencive class 2 properties in the same Williamsbourg.

# Multivariate Analysis

### Talk about some of the relationships you observed in this part of the investigation. Were there features that strengthened each other in terms of looking at your feature(s) of interest?

I noticed a group of points with prices lower than $1000, this looks strange to me, maybe these are mistakes. 
Looks like units of class 2 at average are cheaper and smaller than class 1 and units of class 4 are larger and more expensive.
Properties larger than 4000 feet are mostly commercial or condo, and most small residentials have area less than 5000 feet.
Price distribution varies significantly in different neighborhoods, for example, I notice clusters of lower price class 4 points in Bedford Stuyvesant and Williamsbourg and more expencive class 2 properties in the same Williamsbourg.

### Were there any interesting or surprising interactions between features?

Among the properties built in the beginning of the last century commercial (class 4) properties have higher prices than small residential (class 1) and condos (class 2). On the other hand, commercial properties built in 21st century mostly cost less than residential and condos.

### OPTIONAL: Did you create any models with your dataset? Discuss the strengths and limitations of your model.

------

# Final Plots and Summary

### Plot One

```{r echo=FALSE,warning=FALSE, Plot_one}

ggplot(sales, aes(x = sale.price/1000)) + 
  geom_histogram(bins = 200, fill = 'darkblue') +
  scale_x_log10(breaks = c(1,10,100,1000,10000)) +
  labs(x = "Sale price in $1000",y = "Number of sales",
       title = "Distribution of sale prices on log scale")
```

### Description One

On the log scale distribution looks almost normal with exeption of outliers lower than $1000. Most data are spread between $100 000 and $ 10 000 000 with the mode around $1 000 000. It is interesting that there are a lot more sales just below $1 mln than at $1 mln. I can suppose this is kind of a psycological number or it is connected with tax regulations.

### Plot Two

```{r echo=FALSE, warning=FALSE}
ggplot(filter(sales,price.square.foot >0), aes(price.square.foot)) + 
  geom_density(aes(color = tax.class.at.present),adjust = 2) +
  coord_cartesian(xlim = c(0,3000)) +
  scale_color_brewer(palette = "Dark2") +
  labs(title = "Density function of price for square foot",
       x = "Price per square foot, $",
       y = "Density",
       color = "Tax class")
```

### Description Two

This plot indicates that price for tax class 2C has different distribution - higher mean and variance, and all distribution are skewed and have very long right tale.

### Plot Three

```{r echo=FALSE,warning=FALSE}
ggplot(filter(sales,gross.square.feet > 0,sale.price > 0),
       aes(gross.square.feet/total.units,
           sale.price/total.units)) +
  geom_point(aes(color = tax.class.at.time.of.sale),alpha = 0.3) +
  scale_y_log10() +
  scale_x_log10(minor_breaks = seq(0,1000000,10000),
                breaks = c(1000,10000,100000)) +
  scale_color_brewer(palette = 'Set2') +
  geom_smooth() +
  labs(title = "Relationship between floor area and price of one unit",
       x = "Gross square feet ",
       y = "Sale price, $",
       color = "Tax class")
```

### Description Three

This plot shows the relationship between unit sale price and floor area in different tax classes. Here we can see that properties of class 2 (condominiums and coops) mostly have area less than 1000 square feet and cost less than $1 mln. Commercial properties (class 4) are at average larger and cost more than $1 mln. Smooth layer shows that at general price goes up  with increase in size, but we have a lot of variance due to other variables.

------

# Reflection

This dataset contains information about c. 23000 real estates sold in Brooklyn, NY in 2015, described by 21 variables. Analysing individual variables I've found that main features of interest (sale price, gross square feet, land square feet) have significant proportion of missing values. Some categorical values had more than 20 levels and it made visualisation more difficult. As expected, I found that price and size variables have distributions with very long right tale which made me use log transformation.

After that I explored relationships between price and floor area, neghborhood, building year and tax class. Obviously price is positively correlated with floor area but a lot of variance depends on location. I'm interested in further exploration of reasons behind relationships between price and year built and price and tax class.

Also it would be interesting to build a model for price prediction and find some methods for imputation of missing data. As real estate prices change with time it will be in my opinion the main limitaion to use of the model built on this data.
